---
title: "Real Value Report"
output: 
  officedown::rdocx_document:
    reference_docx: Mission Economics Template.docx
params:
  nominal_value: NULL
  price_year: NULL
  adjusted_year: NULL
  adjusted_value: NULL
  cumulative_inflation: NULL
  avg_inflation: NULL
  year_type: NULL
  value_type: NULL
  deflators_df: NULL
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(officer)
library(flextable)
```

# `r format(Sys.Date(), "%d %B %Y")`

## Summary

Your social value has been adjusted from **£`r params$nominal_value`**
(`r params$price_year` prices) to **£`r params$adjusted_value`**
(`r params$adjusted_year` prices).

### Details

- Calculation basis: `r if (params$year_type == 'fy') 'Financial Years' else 'Calendar Years'` -
- Value type: `r if (params$value_type == 'wellbeing') 'Wellbeing Values' else 'Standard Social Values'` 
- Price change: `r params$cumulative_inflation`% 
- Average annual change: `r params$avg_inflation`%

## Understanding Your Adjusted Value

### Why update my value?

Your updated value more accurately reflects today's economic conditions.
This adjustment is essential for:

1.  Making fair cost-benefit comparisons
2.  Ensuring your social impact isn't undervalued
3.  Meeting H.M. Treasury Green Book requirements

### Methodology

We use the GDP deflator, the Treasury's recommended inflation index because it: 

- Captures economy-wide price changes 
- Is the best measure for public sector and non-market impacts 
- Follows Government guidance for Business Cases

You selected
`r if (params$value_type == 'wellbeing') 'wellbeing' else 'standard'`
values. For these we use the following formula:

`r if (params$value_type == 'wellbeing') {   "Real Value = Nominal Value × (GDP Deflator[Real Year] / GDP Deflator[Nominal Year]) × (GDP per Capita[Real Year] / GDP per Capita[Nominal Year])^1.3" } else {   "Real Value = Nominal Value × (GDP Deflator[Real Year] / GDP Deflator[Nominal Year])" }`

`r if (params$value_type == 'wellbeing') {   "The additional GDP per capita adjustment (with power 1.3) accounts for the diminishing marginal utility of income as society's wealth increases." }`

## Technical Notes

### Sources

-   H.M. Treasury Green
    Book: https://www.gov.uk/government/publications/the-green-book-appraisal-and-evaluation-in-central-government
-   H.M. Treasury (2021) Wellbeing Guidance for
    Appraisal: https://www.gov.uk/government/publications/green-book-supplementary-guidance-wellbeing)
-   H.M. Treasury GDP
    Deflator: https://www.gov.uk/government/collections/gdp-deflators-at-market-prices-and-money-gdp)
-   Office for National Statistics GVA per
    head: https://www.ons.gov.uk/economy/grossdomesticproductgdp/timeseries/ihxw/pn2)

### Version

-   GDP Deflator series (standard social values): HM Treasury Autumn
    Budget, 30th October 2024
-   GVA per capita series (wellbeing values): ONS release, 1st November
    2024
-   Next planned update: When HM Treasury release new GDP deflator
    series ahead of 2025 Spending Review

## Next Steps

1.  Download the complete deflator series (CSV) from the app for batch
    calculations
2.  Consider whether your values need converting to Present Values
3.  Use these adjusted figures in your business case or impact reporting

\newpage

## Appendix: Reference Data

### GDP Deflators and GVA per Head

```{r, echo=FALSE, results='asis'}
df <- params$deflators_df
ft <- flextable(df)
ft <- theme_vanilla(ft)
ft <- autofit(ft)
ft <- delete_part(ft, part = "header")
ft <- add_header_row(ft, values = c("Year", "Calendar Year Deflator", "Financial Year Deflator", "Financial Year Label", "GVA per Head (£)"))
ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

------------------------------------------------------------------------

*Report generated using Get Real*\
*A Mission Economics tool: www.missioneconomics.org*
