---
title: "Get Real Report"
output: 
  officedown::rdocx_document:
    reference_docx: style_reference.docx
params:
  nominal_value: NULL
  price_year: NULL
  adjusted_year: NULL
  adjusted_value: NULL
  cumulative_inflation: NULL
  avg_inflation: NULL
  year_type: NULL
  value_type: NULL
  deflators_df: NULL
---

```{r setup, include=FALSE}
library(officer)
library(flextable)
```

# Get Real Report: `r format(Sys.Date(), "%d %B %Y")`

## Your Value

Your social value of **£`r params$nominal_value`** in **`r params$price_year`** prices is equivalent to **£`r params$adjusted_value`** in 'real terms' (**`r params$adjusted_year`** prices).

Cumulative inflation was **`r params$cumulative_inflation`%** over this period**,** an average of **`r params$avg_inflation`%** inflation per year. The estimates are based on **`r if (params$year_type == 'fy') 'financial years' else 'calendar years'`** and **`r if (params$value_type == 'wellbeing') 'wellbeing values' else 'standard social values'`**.

To validate these estimates, or to estimate social values over multiple years, you can download a CSV file with the deflator series from the Get Real app.

## Why do we need to Get Real?

Proper adjustment of social values is crucial for:

-   **Accurate comparison:** Comparing social benefits to investment costs in 'real terms' is essential for calculating Social Return on Investment (SROI).
-   **Avoiding undervaluation:** High inflation rates over recent years mean that you could be significantly undervaluing your social impacts. For example, values estimated five years ago are around 20% higher.
-   **Credibility:** Maintain credibility with central and local government and other key funders. The H.M. Treasury Green Book provides guidance for economic appraisal and evalaution, used in Business Cases, Impact Assessments and all other public / social value practice. It states that the effects of inflation should be removed with all values expressed in real terms and present values.

## Non-technical Explainer

**Inflation**: Prices increase over time. The rate of this increase is called inflation. When we talk about changes in 'real terms,' we're accounting for inflation to make better comparisons of the actual value of goods and services over time.

**GDP Deflators:** The Treasury recommends using their measures of inflation - the GDP deflator series - to adjust for inflation in social values. The GDP deflator is recommended as a measure of general inflation across the whole economy. You may have heard of other popular measures like the Consumer Price Index (CPI) or Retail Price Index (RPI) - these are also reliable ways to adjust for inflation for the goods and services we buy. But, they are less applicable to estimate the real value of things that we don't buy, like public services. This is why the Treasury and trained social value practitioners use GDP deflators.

**Present Values**: People prefer benefits now rather than later - we call this Social Time Preference. We apply Treasury discount rates to estimate the "present value" of future social costs and benefits. This is a completely separate concept to inflation. Treasury recommend that inflation is removed first, before discounting.

## Technical Guidance

**Inflation Adjustment**: Get Real uses the GDP deflator from the latest Office for Budget Responsibility (OBR) forecasts. It is suitable to estimate values in real terms since the base year is typically the first year of a proposal, or a relatively recent year.

**Short and long-term adjustments:** Get Real is currently suitable for inflating and deflating values up to 2028. For inflation forecasts over longer horizons, Treasury refers practitioners to the OBR Fiscal Sustainability Report.

**General vs specific inflation:** The GDP deflator accounts for general inflation. There can be a case for applying a more specific inflation index, e.g. for construction or information technology. Treasury suggest that relative price effects for specific goods and services can be used in economic appraisal but only where the historical evidence and future expectations support their use. These must be clearly justified.

**Discounting**: We apply the Treasury Social Time Preference Rates (STPR). Over the first thirty years these are 3.5% in real terms and 1.5% for health and wellbeing values. We also apply the incrementally lower long term rates. It is important to convert to real prices first, then apply discounting. Discounting should not be applied retrospectively.

## Version

**GDP Deflators:** Market prices, and money GDP. Outturn data are as at the Quarterly National Accounts from ONS - updated 28 March 2024. Forecast data are consistent with OBR EFO data as at Budget 6 March 2024.

**GVA per head used in WELLBY estimation:** Gross domestic product (Average) per head at market prices - released 10 May 2024.

\newpage

## Annex: GDP deflators and GVA per head

```{r, echo=FALSE, results='asis'}
# Create a nicely formatted flextable
df <- params$deflators_df
ft <- flextable(df)
ft <- theme_vanilla(ft)
ft <- autofit(ft)

# Remove the default header row
ft <- delete_part(ft, part = "header")

# Add a custom header row
ft <- add_header_row(ft, values = c("Year", "Deflator CY", "Deflator FY", "Label FY", "GVA per Capita"))

# Print the table
ft


```
