---
title: "Get Real: Real Values Report"
output: 
  officedown::rdocx_document:
    reference_docx: Mission Economics Template.docx
params:
  nominal_value: NULL
  price_year: NULL
  adjusted_year: NULL
  adjusted_value: NULL
  cumulative_inflation: NULL
  avg_inflation: NULL
  year_type: NULL
  value_type: NULL
  deflators_df: NULL
---

```{r setup, include=FALSE}
library(officer)
library(flextable)
```

# Get Real: Your Social Value in Today's Terms - `r format(Sys.Date(), "%d %B %Y")`

## Your Updated Social Value

Good news! Your social value has been adjusted for inflation:

- Original value: **£`r params$nominal_value`** in **`r params$price_year`** prices
- Updated value: **£`r params$adjusted_value`** in today's terms (**`r params$adjusted_year`** prices)

Over this period, prices increased by **`r params$cumulative_inflation`%** in total, averaging **`r params$avg_inflation`%** per year. We based these calculations on **`r if (params$year_type == 'fy') 'financial years' else 'calendar years'`** and used **`r if (params$value_type == 'wellbeing') 'wellbeing values' else 'standard social values'`**.

Want to check our math or calculate values for multiple years? Download the inflation data (CSV file) from the Get Real app.

## Why Updating Your Social Value Matters

1. **Fair Comparisons:** It lets you accurately compare your social benefits to current investment costs when calculating your Social Return on Investment (SROI).

2. **Avoiding Underestimation:** Recent high inflation means older values might be selling your impact short. For example, a value from five years ago could be about 20% higher today!

3. **Trustworthiness:** Government bodies and funders expect to see up-to-date values. Following official guidance (like the H.M. Treasury Green Book) shows you're doing things by the book.

## Inflation Explained Simply

Imagine you could buy a sandwich for £5 last year, but now it costs £5.50. That 50p increase is inflation in action. When we talk about 'real terms,' we're adjusting for these price changes to make fair comparisons over time.

## How We Update Your Value

We use something called the GDP deflator - it's the government's preferred way to measure overall price changes in the economy. While you might have heard of the Consumer Price Index (CPI), the GDP deflator is better for estimating the real value of things we don't typically buy, like public services.

## For the Technically Curious

- Our app uses forecasts from the Office for Budget Responsibility (OBR) to estimate values up to 2028.
- For longer-term projections, check out the OBR Fiscal Sustainability Report.
- Sometimes, specific inflation rates (e.g., for construction projects) can be used, but this needs solid justification.

## Data Sources

- GDP Deflators: Based on ONS Quarterly National Accounts (updated 28 March 2024) and OBR forecasts from the 6 March 2024 Budget.
- Wellbeing Calculations: Use GDP per head figures released on 10 May 2024.

Need more details? Feel free to ask!

\newpage

## Annex: GDP deflators and GVA per head

```{r, echo=FALSE, results='asis'}
# Create a nicely formatted flextable
df <- params$deflators_df
ft <- flextable(df)
ft <- theme_vanilla(ft)
ft <- autofit(ft)

# Remove the default header row
ft <- delete_part(ft, part = "header")

# Add a custom header row
ft <- add_header_row(ft, values = c("Year", "Deflator CY", "Deflator FY", "Label FY", "GVA per Capita"))

# Print the table
ft


```
